---
title: Presidential Election-night Forecasts
date: "`r paste('UPDATED', as.character(Sys.time(), format='%B %e, %Y %I:%M %p %Z'))`"
output:
    html_document:
        theme: readable
        css: styles_app.css
        fig.width: 9
        fig.height: 6
        self_contained: false
        mathjax: null
runtime: shiny
---


```{r include=F}
library(tidyverse)
library(formattable)
library(shiny)
library(r2d3)

knitr::opts_chunk$set(echo=F, dev.args=list(bg="transparent"), fig.retina=2)
options(r2d3.shadow=F)
transp = theme(panel.background = element_rect(fill = "transparent", colour = NA),
               plot.background = element_rect(fill = "transparent", colour = NA))

states = suppressMessages(read_csv("data/state_ev.csv")) %>%
    select(state, ev=ev.2020) %>%
    left_join(suppressMessages(read_csv("data/state-data.csv")) %>%
                  select(abbr, state),
              by="state") %>%
    arrange(abbr)
states$state[8] = "D.C."
state_ev = set_names(states$ev, states$abbr)
state_names = set_names(states$state, states$abbr)
ap_states = set_names(
    c("Alaska","Ala.", "Ariz.","Ark.","Calif.","Colo.","Conn.","D.C.","Del.",
      "Fla.","Ga.","Hawaii","Idaho","Ill.","Ind.","Iowa","Kan.","Ky.","La.",
      "Mass.","Md.","Maine","Mich.","Minn.","Mo.","Miss.","Mont.","N.C.","N.D.",
      "Neb.","N.H.","N.J.","N.M.","Nev.","N.Y.","Ohio","Okla.","Ore.","Pa.",
      "R.I.","S.C.","S.D.","Tenn.","Texas","Utah","Va.","Vt.","Wash.","Wis.",
      "W.Va.","Wyo."), 
    states$abbr)

dem_states = c("DC", "MA", "HI", "VT", "MD", "RI", "CA", "NY", "DE", "CT", 
               "NJ", "WA", "OR", "IL", "CO")
gop_states = c("WY", "WV", "ND", "ID", "OK", "AR", "UT", "AL", "TN", "KY", 
               "SD", "MS")
close_states = setdiff(states$abbr, c(dem_states, gop_states))

gop_ev = sum(state_ev[gop_states], na.rm=T)
dem_ev = sum(state_ev[dem_states], na.rm=T)

source("R/forecast_update.R")
draws = get_draws()
base_pr = colMeans(select(draws$mine, -draw:-natl) > 0.5)
st_min = apply(select(draws$mine, -draw:-natl), 2, min)
st_max = apply(select(draws$mine, -draw:-natl), 2, max)
```

```{r}
dr_keep = reactive({
    req(input$model)
    filter_ranges(draws[[str_to_lower(input$model)]], close_states, input) 
}) 

renderD3({
    probs = colMeans(select(dr_keep(), -draw:-natl) > 0.5)
    plot_d = states %>% 
        mutate(prob = probs)
    plot_d$state[plot_d$abbr == "DC"] = "District of Columbia"
    
    r2d3(plot_d, script="map.js", 
         dependencies=c("usa.js", "https://d3js.org/topojson.v2.min.js"),
         container="div", elementId="map")
})

renderUI({
    dr = dr_keep();
    probs = colMeans(select(dr, -draw:-natl) > 0.5)
    ci = round(quantile(dr$ev, c(0.05, 0.95)))
    pr_win = round(100*mean(dr$ev >= 270))
    if (pr_win == "100" & sum(dr$ev < 270) > 0)
        pr_win = ">99"
    
    div(
        strong("Joe Biden", style="color: #56e"), "secures",
        strong(sum(state_ev[probs >= 0.999])), "electoral votes, to",
        strong("Donald Trump", style="color: #e65", .noWS="after"), "’s",
        strong(sum(state_ev[probs <= 0.001])), "votes.",
        strong("Biden", style="color: #56e"), "has a",
        strong(pr_win, "in 100"),
        "in chance of winning the election.",
        "He is expected to win a total of ",
        strong(ci[1], "–", ci[2]),
        "electoral votes.",
    class="banner")
})

renderPlot({
    evs = dr_keep()$ev
    colors = c("#ee6655", "#5566ee")[1 + (evs >= 270)]
    ggplot(NULL, aes(x=evs, fill=colors)) + 
        geom_histogram(binwidth=1) +
        geom_vline(xintercept=270, lty="dashed") +
        scale_fill_identity() +
        guides(fill=F) +
        coord_cartesian(expand=F, xlim=c(0.5*dem_ev, 538-0.5*gop_ev)) +
        labs(y=NULL, x="Electoral votes for Biden") +
        transp +
        theme(panel.grid=element_blank(),
              axis.line.x.bottom=element_line(),
              axis.text.y=element_blank(),
              axis.ticks.y=element_blank())
}, height=220, bg="transparent", res=92)
```


# State Calls
Use the sliders below to set Biden's share of the two-party vote in every state.

The model starts with thousands of possible election scenarios, or "draws." The
page will update depending on your selections by throwing away the draws that
don't match what you've selected. You need at least 50 draws to have reliable
estimates; the more the better.

Right now you are using  __`r renderText(scales::comma(nrow(dr_keep())))` draws__.

```{r}
    selectInput("model", "Forecasting model", c("Mine", "Economist"), selectize=F)
```

### Lean Trump
```{r}
sl_gop = setdiff(states$abbr[base_pr < 0.3], gop_states)
sl_gop = sl_gop[order(base_pr[sl_gop])]
div(renderUI({
    map(sl_gop, function(s) {
        sliderInput(s, label=state_names[s], min=25, max=65,
                    value=round(c(st_min[s]*97, st_max[s]*103), 1), step=0.1,
                    ticks=F, width="15em", dragRange=F) 
    })
}), class="state_sliders gop")

```

### Tossups
```{r}
sl_toss = states$abbr[base_pr >= 0.3 & base_pr <= 0.7]
sl_toss = sl_toss[order(base_pr[sl_toss])]
div(renderUI({
    map(sl_toss, function(s) {
        sliderInput(s, label=state_names[s], min=30, max=70,
                    value=round(c(st_min[s]*97, st_max[s]*103), 1), step=0.1,
                    ticks=F, width="14em", dragRange=F) 
    })
}), class="state_sliders toss")

```


### Lean Biden
```{r}
sl_dem = setdiff(states$abbr[base_pr > 0.7], dem_states)
sl_dem = sl_dem[order(base_pr[sl_dem])]
div(renderUI({
    map(sl_dem, function(s) {
        sliderInput(s, label=state_names[s], min=35, max=75,
                    value=round(c(st_min[s]*97, st_max[s]*103), 1), step=0.1,
                    ticks=F, width="14em", dragRange=F) 
    })
}), class="state_sliders dem")

```

# State-by-State Odds
```{r}
renderFormattable({
    probs = colMeans(select(dr_keep(), -draw:-natl) > 0.5)
    called = abs(0.5-probs) >= 0.495
    
    states %>% 
        select(State=state, EVs=ev) %>% 
        filter(!called) %>%
        mutate(`Biden's chances` = percent(probs[!called], 0),
               `Trump's chances` = 1 - `Biden's chances`) %>%
        arrange(`Trump's chances`) %>%
    formattable(list(
        `Biden's chances` = color_tile("#ffffff00", "#5566ee"),
        `Trump's chances` = color_tile("#ffffff00", "#ee6655")
    ))
})
```
