---
title: Presidential Election-night Forecasts
date: "`r paste('November 3, 2020,', as.character(Sys.time(), format='%I:%M %p %Z'))`"
author: Cory McCartan
output:
    html_document:
        theme: readable
        css: styles.css
        fig.width: 9
        fig.height: 6
        self_contained: false
        mathjax: null
---

```{r include=F}
library(usmap)
library(ggplot2)
library(formattable)

knitr::opts_chunk$set(echo=F, dev.args=list(bg="transparent"), fig.retina=2)
transp = theme(panel.background = element_rect(fill = "transparent", colour = NA),
               plot.background = element_rect(fill = "transparent", colour = NA))
```

## National Overview
```{r fig.width=9, fig.height=6}
probs = calls
state_names = set_names(states$state, states$abbr)
pred_states = state_names[colnames(preds$states)]
est_probs = apply(preds$states, 2, function(x) weighted.mean(x > 0.5, preds$w))
probs[names(est_probs)] = est_probs

us_map("states") %>%
    mutate(prob = probs[abbr],
           called = as.numeric(prob >= 0.999 | prob <= 0.001)) %>%
ggplot(aes(x, y, group=group, fill=prob)) + 
    geom_polygon(color=NA, size=0) + 
    geom_polygon(aes(size=called, color=called), fill=NA) + 
    scale_fill_gradient2(low="#ee6655", high="#5566ee", mid="#f0f0f0", 
                         midpoint=0.5, limits=c(0, 1), na.value="#c0c0c0") +
    scale_size_continuous(range=c(0.2, 0.65)) +
    scale_color_gradient(low="#444444", high="#000000") +
    guides(fill=F, color=F, size=F) +
    coord_fixed() + 
    theme_void() +
    transp
```

```{r}
dem_ev = sum(state_ev[calls==1], na.rm=T)
gop_ev = sum(state_ev[calls==0], na.rm=T)
distr_ev = dem_ev + as.numeric((preds$states[preds$rs,] > 0.5) %*% state_ev[colnames(preds$states)])
```

<div class="banner">
<b style="color: #56e">Joe Biden</b> has won <b>`r dem_ev`</b> electoral votes, 
to <b style="color: #e65">Donald Trump</b>'s <b>`r gop_ev`</b> votes. <br />
<b style="color: #56e">Biden</b> has a <b>`r round(100*mean(distr_ev >= 270))` in 100</b>
chance of winning the election.
</div>

```{r fig.width=9, fig.height=3}
ev_hist = function(evs) {
    ggplot(NULL, aes(x=evs, fill= evs>=270)) + 
        geom_histogram(binwidth=1) +
        geom_vline(xintercept=270, lty="dashed") +
        scale_fill_manual(values=c("#ee6655", "#5566ee")) +
        guides(fill=F) +
        coord_cartesian(expand=F, xlim=c(0.5*dem_ev, 538-0.5*gop_ev)) +
        labs(y=NULL, x="Electoral votes for Biden") +
        transp +
        theme(panel.grid=element_blank(),
              axis.line.x.bottom=element_line(),
              axis.text.y=element_blank(),
              axis.ticks.y=element_blank())
}

ev_hist(distr_ev)
```

# Key States
The remaining uncalled states:
```{r}
tibble(
    State = pred_states,
    `Estimated Biden vote` = percent(apply(preds$states, 2, function(x) weighted.mean(x, preds$w))),
    `Chance Biden wins` = percent(est_probs, digits=0)
) %>%
    arrange(desc(`Chance Biden wins`)) %>%
    formattable(list(
        `Chance Biden wins` = color_tile(min.color="#ee6655cc", max.color="#5566eecc"),
        `Estimated Biden vote` = normalize_bar("gray", min=0.4, max=0.6)
        ))
```

The closest state:

```{r fig.height=4, fig.width=8}
idx_min = which.min((est_probs-0.5)^2)
qplot(preds$states[,idx_min], bins=40,
      main=pred_states[idx_min], xlab="Biden vote share") +
    scale_x_continuous(labels=percent) +
    geom_vline(xintercept=0.5, lty="dashed") +
    coord_cartesian(expand=F) +
    transp +
    theme(panel.grid=element_blank(),
          axis.line.x.bottom=element_line(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())
```

# Scenarios {.tabset}

```{r}
dem_called = as.character(na.omit(names(calls)[calls==1]))
gop_called = as.character(na.omit(names(calls)[calls==0]))
to_call = names(calls)[is.na(calls)]

filtered = filter_wins(draws, dem_called, gop_called)

hypotheticals = map_dfr(to_call, function(s) {
    if_dem = filter_wins(filtered, dem_win=s, gop_win=gop_called) %>% 
        imap_dfr(~ tibble(State=state_names[s], EVs=state_ev[s], 
                          win="If Biden wins", forecast=.y, 
                          n=nrow(.x), prob=mean(.x$ev >= 270)))
    if_gop = filter_wins(filtered, dem_win=dem_called, gop_win=s) %>% 
        imap_dfr(~ tibble(State=state_names[s], EVs=state_ev[s],
                          win="If Trump wins", forecast=.y, 
                          n=nrow(.x), prob=mean(.x$ev >= 270)))
    
    bind_rows(if_dem, if_gop) %>%
        mutate(prob = percent(prob))
})

hyp_table = function(cast) {
    hypotheticals %>%
        filter(n > 30, forecast==cast)  %>%
        select(-n, -forecast) %>%
        pivot_wider(names_from=win, values_from=prob) %>%
        arrange(desc(abs(`If Biden wins` - `If Trump wins`))) %>%
        select(State, EVs, `If Trump wins`, `If Biden wins`) %>%
    formattable(list(
        `If Biden wins` = color_tile("#ffffff00", "#5566ee"),
        `If Trump wins` = color_tile("#ffffff00", "#ee6655")
    ))
}
```

## Main Forecast

```{r fig.width=9, fig.height=3}
ev_hist(filtered$mine$ev)
```

Based on the states that have been called, the main forecast estimates that
<b style="color: #56e">Biden</b> has a <b>`r round(100*mean(filtered$mine$ev >= 270))` in 100</b>
chance of winning the election.

This table shows Biden's chances of winning the election if he or Trump wins
specific remaining states.
```{r}
hyp_table("mine")
```


## _Economist_ Forecast

```{r fig.width=9, fig.height=3}
ev_hist(filtered$economist$ev)
```

Based on the states that have been called, the _Economist_ forecasts that
<b style="color: #56e">Biden</b> has a <b>`r round(100*mean(filtered$economist$ev >= 270))` in 100</b>
chance of winning the election.

This table shows Biden's chances of winning the election if he or Trump wins
specific remaining states.
```{r}
hyp_table("economist")
```
